How cryptic is cryptic diversity? Machine learning approaches to classifying morphological variation in \textit{Emys marmorata} (Testudinoidea, Emydidae).

### 2014-11-05 doc/draft_turtle.tex
permutation test write up

In classification studies, like this one, a common metric of performance is area under the reciver operating characteristic curve (AUC). AUC is an estimate of the relationship between the false positive and true positive rates, as opposed to just the true positive rate (accuracy). This relationship is especially useful in cases where misclassification needs to be minimized just as much as accurate classification, as in this study. AUC ranges between 0.5 and 1, with 0.5 indicating classification no better than random and 1 indicating perfect classification \citep{Hastie2009}.

@@

>>>>

In classification studies, like this one, a common metric of performance is area under the receiver operating characteristic curve (AUC). AUC is an estimate of the relationship between the false positive and true positive rates, as opposed to just the true positive rate (accuracy). This relationship is especially useful in cases where misclassification needs to be minimized just as much as accurate classification, as in this study. AUC ranges between 0.5 and 1, with 0.5 indicating classification no better than random and 1 indicating perfect classification \citep{Hastie2009}.

@@

It is common for some out of sample observations to be misclassified. This misclassification may be due to the model not accurately representing shape variance, systematic differences between the training and test sets, or systematic differences between the accurately and inaccurately classified samples. Testing and training sets are determined completely at random within each class and with respect to shape. Results are not effected by changes in testing or training set assignment.

To determine if there were systematic differences between the correctly and incorrectly classified samples, the multivariate centroids of the correct and incorrect groups were compared to what would be expected by random chance. The group labels were permuted and the difference between the new centroids was calculated. This was repeated 1000 times. The number of permutations less than the empirical difference were counted and divided by 1000, giving a \textit{p}-value. Significant results indicate that correctly and incorrectly classified specimens are systematically different.
